\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{amsfonts}
\usepackage{pseudocode}
\usepackage{blkarray}
\usepackage{setspace}
\usepackage{etoolbox}
\newcommand{\blind}{0}



\addtolength{\oddsidemargin}{-.75in}%
\addtolength{\evensidemargin}{-.75in}%
\addtolength{\textwidth}{1.5in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\graphicspath{{figures_tables/}}

\begin{document}
\newtoggle{thesis}
\togglefalse{thesis}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}
  
  
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    
    \if0\blind
  {
    \title{\bf A GPU accelerated nonparametric shrinkage model for high-dimensional data}
    \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
    and\\
    Jarad Niemi\footnotemark[1]}
      
      
    \maketitle
  } \fi
  
  \if1\blind
  {
    \bigskip
    \bigskip
    \bigskip
    \begin{center}
    {\LARGE\bf A GPU accelerated nonparametric shrinkage model for high-dimensional data}
    \end{center}
    \medskip
  } \fi
  
  \bigskip
  \begin{abstract}
  In high-dimensional settings where traditional shrinkage priors are commonly applied, we consider an alternative nonparametric Bayesian approach based on a Dirichlet process mixture. This highly-flexible approach ``learns" the latent distribution of the random effects to allow for efficient shrinkage toward that distribution. In order to make our approach computationally feasible, we take advantage of the massively parallel architecture of the GPU. We discuss the details of the model and specifics of our implementation of the Gibbs sampler. We use gene expression profiling experiments as a motivating example. Through simulations we study the how the computational time requirements scale with the data. We illustrate our method with a differential expression analysis on a blocked RNA-seq experiment.
  \end{abstract}
  
  \noindent%
  {\it Keywords:}  Bayesian nonparametrics, GPU, high-dimensional, gene expression, Dirichlet process, Gibbs sampling
  
  \spacingset{1.45}
  
  \include{chapter2}

  
  \bigskip
  \begin{center}
  {\large\bf SUPPLEMENTAL MATERIALS}
  \end{center}
  
  \begin{description}
  
  \item[Title:] Brief description. (file type)
  
  \item[R-package for  MYNEW routine:] R-package ÒMYNEWÓ containing code to perform the diagnostic methods described in the article. The package also contains all datasets used as examples in the article. (GNU zipped tar file)
  
  \item[HIV data set:] Data set used in the illustration of MYNEW method in Section~ 3.2. (.txt file)
  
  \end{description}
  \bibliographystyle{apalike}
  \bibliography{computation}
  
  \end{document} 